Malacology: A Programmable Storage System

### 2016-10-21 paper/paper.tex
paper: reduce captions and cut figures in half
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}
\caption{Scalable storage systems consist of storage daemons which store data
fragments, monitor daemons (M) that maintain cluster state, and service-specific
daemons such as file system metadata servers. Malacology enables the
programmability of internal abstractions (indicated by bold arrows) in order to
re-use and compose existing subsystems.  With Malacology, we add two new
services, ZLog and Mantle, that sit alongside the traditional user-facing APIs
(file, block, and object) in the storage stack.  \label{fig:overview}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
interface commit.}
\label{fig:propdelay}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-balancer-behaviors.png}
\caption{The load balancing interface lets administrators control the migration
of sequencers in environments with multiple ZLog logs. For this setup (3
sequencers and 3 servers), CephFS and Mantle migrate load and show improved
throughput over the setup where all sequencers are co-located on the same
server (``No Balancing").  Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify the performance improvement of
the stable phase at time 100 seconds and Section~\ref{sec:feature-backoff}
examines the growth phase at time 0 to 60 seconds.
}\label{fig:mantle-balancer-behaviors}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-balancer-performance.png}
\caption{Despite using different balancing modes, the CephFS balancers have the
same performance because they end up migrating the sequencers in the same way.
The Mantle bar programs a balancer designed for sequencers into the service
using the load balancing interface. This is the same workload/setup as
Figure~\ref{fig:mantle-balancer-behaviors}.}\label{fig:mantle-balancer-performance}
\end{figure}

@@

Using Mantle, the administrator can put the load balancing service into ``proxy
mode" or ``client mode". In proxy mode one server receives all requests and
farms off the requests to slave servers; the slaves servers do the actual tail
finding operation. In client mode, clients interact directly with the server
that has their sequencer.  These modes are illustrated in
Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
co-located on one physical server -- performance for that mode is shown by the
No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
Mode" in the lower left shows how clients for sequencer 2 are re-directed to
server B. The two ``Proxy mode" diagrams on the right show how clients continue
sending reqeusts to server A even though sequencer 2 has migrated to server B
(the difference between ``Proxy Mode (Half)" and ``Proxy Mode (Full)" is
discussed in the next section).

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-modes.png}
\caption{The client and proxy balancing modes work well for migrating
sequencers to alleviate overloaded servers. Client mode migrates the sequencer
and clients start sending requests to the new server. Proxy mode migrates the
sequencer but clients continue sending their requests to the first server.
``Half" is the size of the migration units (discussed in
Section~\ref{sec:feature-migration-units}).}\label{fig:mantle-modes}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-mode-behavior.png}
\caption{For this 2 sequencer/2 server setup, the performance of proxy mode
achieves the highest throughput but at the cost of lower throughput for one of
the sequencers. Client mode is more fair but results in lower cluster
throughput; the difference in throughput amongst sequencers can be attributed
to one of the servers doing cache coherence.
}\label{fig:mantle-mode-behavior}
\end{figure}

@@

Another factor that affects performance in this environment is ``how much" load
is on each server; these experiments quantify that effect by programming
policies into the load balancing interface to control the amount of load to
migrate. We defined a  migration unit as the amount of load to try and migrate.
For example, it is useful to express migration units as a fraction of the total
load.  Expressing this heuristic is not easily achievable with outward facing
tunable parameters (i.e. system knobs) and so the original CephFS balancers do
not have a migration unit tunable.  With programmability, migration units can
be expressed quite easily. To force the load balancer into the Proxy Mode
(Half) scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to
proxy mode with migration units of half the load on the current server we can
use the Mantle policy:\\

\noindent \texttt{ targets[\#mds+1] = mds[whoami]["load"]/2 }\\

This code snippet uses globally defined variables and tables from the Mantle
API to send half of the load on the current server (whoami) to the next ranked
server (\#mds + 1); the \texttt{targets} array is a globally defined table that
the balancer uses to do the migrations. Alternatively, to set the mode to the
``Proxy (Full)"` setup in Figure~\ref{fig:mantle-modes} we use:\\

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-mode-performance.png}
\caption{The performance of the different modes using various migration units,
indicated by parentheses, shows almost a 2\(\times\) improvement in the best
case.  }\label{fig:mantle-mode-performance}
\end{figure}

Figure~\ref{fig:mantle-mode-performance} shows the performance of the modes
using different migration units. Recall that this setup only has 2 sequencers
and 2 servers, so performance may be different at scale. Even so, it is clear
that client mode does not perform as well for read-heavy workloads. We
even see a throughput improvement when migrating all load off the first server,
leaving the first server to do administrative tasks (this is common in the
metadata cluster because the first server does a lot of the cache coherence
work) while the second server does all the client work. Proxy mode does the
best in both cases and shows large performance gains when completely decoupling
client request handling and operation processing in Proxy Mode (Full).

The parameter that controls the migration units helps the administrator control
the sequencer co-location or distribution across the cluster. This trade-off
was explored extensively in the Mantle paper but the experiments we present
here are indicative of an even richer set of states to explore.

@@

We control the conditions for migration using \texttt{when()}. \texttt{when()}
is part the Mantle API and the callback expects a return value of true or
false, which corresponds to whether to migrate or not. For the Mantle curve in
Figure~\ref{fig:mantle-balancer-behaviors} we set the \texttt{when()} policy to
wait for load on the receiving server to fall below a threshold. At the
beginning of the run the load of the receiving server spikes because requests
for cache-coherence start as the sequencers are created. It takes 60 seconds
for this load to drop to an acceptable level and it is at this point that the
migration starts.

The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} also takes longer
to reach peak throughput.  This conservative decision making is intentional
because we want the policy wait to see how migrations affect the system before
proceeding. The Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does
a migration right before 50 seconds, realizes that there is a third underloaded
server, and does another migration.

>>>>

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}
\caption{Scalable storage systems have storage daemons which store data,
monitor daemons (M) that maintain cluster state, and service-specific daemons
({\it e.g.}, file system metadata servers). Malacology enables the programmability of
internal abstractions (bold arrows) to re-use and compose
existing subsystems.  With Malacology, we add new services, ZLog and
Mantle, that sit alongside traditional user-facing APIs (file, block,
object).}\label{fig:overview}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=75mm,trim={1 4 4 1.3cm},clip]{figures/iface-update-delay.pdf}
\caption{Cluster-wide interface update latency, excluding the Paxos proposal cost for
interface commit.}
\label{fig:propdelay}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-balancer-behaviors.png}
\caption{CephFS/Mantle load balancing have better throughput than co-locating
all sequencers on the same server.  Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify this improvement;
Section~\ref{sec:feature-backoff} examines the migration at 0-60 seconds.
}\label{fig:mantle-balancer-behaviors}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-balancer-performance.png}
\caption{All CephFS balancing modes have the same performance Mantle uses a
balancer designed for sequencers.}\label{fig:mantle-balancer-performance}
\end{figure}

@@

Using Mantle, the administrator can put the load balancing service into ``proxy
mode" or ``client mode". In proxy mode one server receives all requests and
farms off the requests to slave servers; the slaves servers do the actual tail
finding operation. In client mode, clients interact directly with the server
that has their sequencer.  These modes are illustrated in
Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
co-located on one physical server -- performance for that mode is shown by the
No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
Mode" in the lower left shows how clients for sequencer 2 are re-directed to
server B. The two ``Proxy mode" diagrams on the right show how clients continue
sending reqeusts to server A even though sequencer 2 has migrated to server B
(the meaning of ``(Half)" is discussed in the next section).

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-modes.png}
\caption{ In client mode clients sending requests to the server that houses
their sequencer. In proxy mode clients continue sending their requests to the
first server.  }\label{fig:mantle-modes}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-mode-behavior.png}
\caption{The performance of proxy mode achieves the highest throughput but at
the cost of lower throughput for one of the sequencers. Client mode is more
fair but results in lower cluster throughput.
}\label{fig:mantle-mode-behavior}
\end{figure}

@@

Another factor that affects performance in this environment is ``how much" load
is on each server; these experiments quantify that effect by programming
policies into the load balancing interface to control the amount of load to
migrate. We define migration unit as the amount of load to try and migrate.
For example, it is useful to express migration units as a fraction of the total
load.  Expressing this heuristic is not easily achievable with outward facing
tunable parameters (i.e. system knobs) and so the original CephFS balancers do
not have a migration unit tunable.  With Mantle's programmable interface, we
force the load balancer into the Proxy Mode (Full) scenario in
Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy mode with
migration units equal to the entire load on the current server we can use the Mantle
policy:\\

@@

This code snippet uses globally defined variables and tables from the Mantle
API to send half of the load on the current server (whoami) to the next ranked
server (\#mds + 1); the \texttt{targets} array is a globally defined table that
the balancer uses to do the migrations. Alternatively, to set the mode to the
``Proxy (Half)"` setup in Figure~\ref{fig:mantle-modes} we can divide the right
hand side by 2.

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-mode-performance.png}
\caption{The performance of the different modes using various migration units
shows almost a 2\(\times\) improvement in the best case.
}\label{fig:mantle-mode-performance}
\end{figure}

Figure~\ref{fig:mantle-mode-performance} shows the performance of the modes
using different migration units. Recall that this setup only has 2 sequencers
and 2 servers, so performance may be different at scale. Even so, it is clear
that client mode does not perform as well for read-heavy workloads. We even see
a throughput improvement when migrating all load off the first server, leaving
the first server to do administrative tasks (this is common in the metadata
cluster because the first server does a lot of the cache coherence work) while
the second server does all the client work. Proxy mode does the best in both
cases and shows large performance gains when completely decoupling client
request handling and operation processing in Proxy Mode (Full).  The parameter
that controls the migration units helps the administrator control the sequencer
co-location or distribution across the cluster. This trade-off was explored
extensively in the Mantle paper but the experiments we present here are
indicative of an even richer set of states to explore.

@@

We control the conditions for migration using \texttt{when()}. \texttt{when()}
is part the Mantle API and the callback expects a return value of true or
false, which corresponds to whether to migrate or not. For the Mantle curve in
Figure~\ref{fig:mantle-balancer-behaviors} we set the \texttt{when()} policy to
wait for load on the receiving server to fall below a threshold. At the
beginning of the run the load of the receiving server spikes because requests
for cache-coherence start as the sequencers are created. It takes 60 seconds
for this load to drop to an acceptable level and it is at this point that the
migration starts.  The Mantle curve in
Figure~\ref{fig:mantle-balancer-behaviors} also takes longer to reach peak
throughput.  This conservative decision making is intentional because we want
the policy wait to see how migrations affect the system before proceeding. The
Mantle curve in Figure~\ref{fig:mantle-balancer-behaviors} does a migration
right before 50 seconds, realizes that there is a third underloaded server, and
does another migration.
